---
title: "Two Parameters"
author: Faran Abbas
format: html
execute: 
  echo: false
---


```{r}
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(performance)
library(dplyr)
library(marginaleffects)
```




Height is an important physical characteristic influenced by factors such as age and sex. This analysis uses data from the NHANES survey collected by the CDC to explore how sex affects height in young adults aged 18 to 27.

$$Y_i = \beta_0 + \beta_1 X_{1i} + \beta_2 X_{2i} + \cdots + \beta_k X_{ki} + \epsilon_i$$
$$\mathbf{Y} = \mathbf{X}\boldsymbol{\beta} + \boldsymbol{\epsilon}$$
$$\epsilon_i \sim \mathcal{N}(0, \sigma^2)$$

**Weakness**

"The model assumes that error terms are independently and identically distributed with constant variance (homoscedasticity), but if this assumption is violated due to heteroscedasticity or autocorrelation, the parameter estimates may be inefficient and statistical inferences unreliable."


```{r}
library(NHANES)
library(tidymodels)
library(dplyr)

set.seed(10)

# Recreate the dataset x
x <- NHANES |>
  filter(Age >= 18 & Age <= 27) |>
  select(Height, Gender) |>
  drop_na() |>
  slice_sample(n = 50)

# Then fit the model
linear_reg() |> 
  fit(Height ~ Gender, data = x) |>
  tidy(conf.int = TRUE)
```


$$\hat{\text{height}} = 162.18 + 15.94 \times I(\text{sex} = \text{Male})$$


The fitted regression model is:
$$\hat{\text{height}} = 162.18 + 15.94 \times I(\text{sex} = \text{Male})$$

where $\text{sexMale}$ is an indicator variable equal to 1 for males and 0 for females.

Model Estimation
```{r}
#| cache: true
library(NHANES)
library(tidymodels)
library(dplyr)

set.seed(10)

x <- NHANES |>
  filter(Age >= 18 & Age <= 27) |>
  select(height = Height, sex = Gender) |>
  drop_na() |>
  slice_sample(n = 50)

fit_height <- linear_reg() |> 
  fit(height ~ sex, data = x)
```

Model Results
```{r}
#| echo: false
library(knitr)
fit_height |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  kable(digits = 2, 
        col.names = c("Term", "Estimate", "95% CI Lower", "95% CI Upper"))

```


```{r}
library(marginaleffects)
predictions(fit_height)
avg_predictions(fit_height)
avg_predictions(fit_height, by = "sex")
plot_predictions(fit_height, condition = "sex")

plot_predictions(fit_height, condition = "sex", draw = FALSE)
```

```{r}
#| fig-width: 8
#| fig-height: 5
plot_predictions(fit_height, condition = "sex") +
  geom_point(data = x, aes(x = sex, y = height), 
             alpha = 0.5, size = 2, position = position_jitter(width = 0.2)) +
  labs(title = "Height by Sex in Young Adults (Ages 18-27)",
       subtitle = "Males are, on average, 15.94 cm taller than females.",
       x = "Sex",
       y = "Height (cm)",
       caption = "Data source: NHANES survey, CDC") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    plot.caption = element_text(hjust = 0, size = 10),
    axis.title = element_text(face = "bold")
  )
```


**Project Summary**
We model height, measured in centimeters, as a linear function of sex (male or female) to examine potential differences in average height between gender groups. On average, males are 15.94 cm taller than females, with a 95% confidence interval of [11.36, 20.52] cm.